Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations" \\ \textit{The~American~Naturalist}

### 2021-12-10 doc/ResponseLetter-R1.tex

\begin{quote}
	{\itshape ...}
\end{quote}

\noindent Apart from that, I have only minor comments, that I list below (roughly in decreasing order of importance):
\bigskip

\noindent -- Model: lines 114-195
\begin{itemize}
	\item The matrices H are mentioned before they are introduced. (Therefore, the parenthesis in lines 123-124 cannot be understood by the reader). And then are only defined in the following section	
	\begin{itemize}
		\item It would make sense to reorder the model presentation by presenting the building blocks (arguably U, F and H matrices, but I'll come back to H in the next point), what they do, what they are for; then provide some properties (F independent from selfing/crossing, H from stage, etc ...) and then go into the details of their construction
	\end{itemize}
	\item The main issue I have here, is that H is introduced, but then disappears from the output (see equations 12 and 13 for instance that use H but it does not appear, or the general output eq 9)
	\begin{itemize}
		\item This suggests combining F and H directly into an object.
	\end{itemize}
\end{itemize}

\begin{itemize}
	\item More generally, the matrix formulations of the model may be a bit hard to read for the general reader of AmNat. Getting rid of H (or to the contrary, using it all the way, thus simplifying eq 12 and 13 for instance) and not giving a name to matrix W (which is just a bunch of 0s 1s and 1/2s) are steps that can make it easier for the reader to follow.
\begin{quote}
	{\itshape Response to the above 5 points: As per the reviewers (and the editors) suggestion, we have moved a lot of the derivations into the supplementary materials, and now no longer use H or W in the main text. In doing so, we hope to shift the focus from mathematical derivations, which are probably unhelpful to many readers, towards explaining the end result instead. }
\end{quote}

	\item It is quite confusing that HX is written as a constant (eq 7) whilst it depends on the frequencies and that HS is written as dependent (eq6) whilst it is constant
	\begin{quote}
	{\itshape That's a mistake on our part, apologies.}
\end{quote}
	\item As the building blocks are given expressions (eq 10 and 11 for instance) it would be nice to mention the fact that the elements in these will be characterize later (otherwise the reader thinks she/he should already know what the expression of, say, UsAA is)
\end{itemize}

\begin{quote}
	{\itshape As suggested by this reviewer and by the editor, we have now moved up the two stage example, so that we show the reader what these matrices can look like immediately after introducing them. Again shifting the focus from mathematical derivations to an explicit example, thereby hopefully improving the manuscripts' accessibility/readability.  }
\end{quote}
\noindent -- Line 19: I find it weird to write that the "process of adaptation can be impeded by ...changing environmental conditions".

@@

\noindent -- Line 30: "fitness trade-offs ... provide ...mechanism for the maintenance of genetic variation". This requires a few words of explanation. Is that a general statement about (fitness) trade-offs? or is it actually particular to SA trade-offs or to trade-offs between fixed heterogeneity components?
\begin{itemize}
	\item For instance, a fitness trade off between fertility at 3 years and fertility at 6 years will have no positive effect on genetic variance. To the contrary, it may lead the organism to evolve an optimal strategy accounting for overall fitness and the trade-offs, vis-à-vis investments in reproduction at those different stages in the life cycle. And as such this may lead to the selection of individuals with such a strategy and therefore to low genetic variance.
\end{itemize}

\begin{quote}
	{\itshape ... our response}
\end{quote}

\noindent -- Line 264: this is an important point that is a bit unclear here, according to me. Could you clarify that what you call demographic rates are the transitions in the overall model and so include both demography and genetics? Clarify also that the model is not density dependent because it only depends on frequencies ... and maybe relate it to a general equation where A does not depend on n per se but on the q frequencies.

\begin{quote}
	{\itshape ... our response \textcolor{red}{I cannot figure out where in the manuscript this is, I don't think I have the exact R0 version as it was submitted?}}
\end{quote}

### 2021-12-10 doc/Supplements/SupplementaryMaterial.tex
Supplementaries
Edited supplementaries and added a few more answers to the response letter

\newpage{}
\section{Construction of the population projection matrix}
\subsubsection*{Mating and offspring production under partial selfing}
 Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}. That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}. Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion. For the purpose of this article, we also assume that mating is random with respect to stage.

@@

The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map. As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}). Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis: 
\begin{linenomath*}
\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				1 & 1/4 & 0 \\
				0 & 1/2 & 0 \\
				0 & 1/4 & 1 \\
			\end{array} \right).
\end{equation}
\end{linenomath*}
\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively. As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS}) and Eq(\ref{eq:BlkFX}) below).

Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability. In the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).

We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(10)-(13) in the main text. The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in Eq(9), repeated here for convenience, 
\begin{linenomath*}
\begin{equation} \label{eq:Atilde_appendix}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}}.
\end{equation}
\end{linenomath*}

>>>>

\newpage{}
\section{Construction of the population projection matrix}

In this section we will go through the derivation of the matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of partially-selfing simultaneous hermaphrodites model presented in the main text. We start by calculating the frequencies of alleles in the gamete pool. These frequencies are the ingredients for the map of parent genotype to offspring genotype, which itself is a necessary ingredient to create the fertility matrices. Finally putting the fertility components together with the (slightly simpler) survival matrices, we obtain all the ingredients to build the population projection matrix.

\subsubsection*{Mating and offspring production under partial selfing}
 Outcrossing in our model proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}. That is, each individual's genotype determines both the number of ovules produced and the paternal mating success, broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}. Note, however, that by modeling paternal relative mating success rather than pollen/sperm production, we implicitly assume female demographic dominance (i.e., production/transport of male gametes does not limit ovule fertilization and hence population growth), a point we return to in the Discussion. For the purpose of this article, we also assume that mating is random with respect to stage.

@@

The parent-offspring map for reproduction via self-fertilization differs from the outcrossing parent-offspring map. As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection via the male sex function from external factors (e.g., pollinator visitation) relative to outcrossing (\citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}). Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis: 
\begin{linenomath*}
\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				1 & 1/4 & 0 \\
				0 & 1/2 & 0 \\
				0 & 1/4 & 1 \\
			\end{array} \right).
\end{equation}
\end{linenomath*}
\noindent The columns of $\mbf{H}^S(\tilde{\mbf{n}})$ again give the selfed offspring genotype distributions for parental genotypes of $AA$, $Aa$, and $aa$, respectively. As described in the next section, these parent-offspring maps influence the projection matrix by altering the fertility matrices for individuals produced by selfing and outcrossing (see Eq(\ref{eq:BlkFS_appendix}) and Eq(\ref{eq:BlkFX_appendix}) below).

Following previous theory for partially selfing populations \cite[e.g.,][]{Charlesworth2010,JordanConnallon2014,Glemin2021}, we use a 'fixed prior selfing' model where all individuals are assumed to self-fertilize at a constant rate, $C$, prior to receiving outcross pollen, for reasons of analytic tractability. Further down in the Online Supplement, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}).

\subsubsection*{The Population Projection Matrix}
We will now derive the expressions for the survival and fertility matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ as given by Eq(12)-(15) in the main text. The matrices $\mcal{U}^S$, $\mcal{U}^X$, $\mcal{F}^S$, and $\mcal{F}^X$ are the ingredients of the population projection matrix, $\tilde{\mbf{A}}$, as given in Eq(9), repeated here for convenience, 
\begin{linenomath*}
\begin{equation} \label{eq:Atilde_appendix}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}}.
\end{equation}
\end{linenomath*}
